Kaggle-in-class Data Challenges Can Boost Student Learning

### 2020-07-08 paper-kaggle.Rmd
Fixing reviewers comments

The reason for this strategy was first to motivate each of the students to think about modelling and be actively engaged in the competitions through individual submission. The lecturer allowed participants to create groups towards the end of the competition to illustrate the advantages of group work and ensemble models. Another reason for this stratagem was the university policy, requiring a strategy to assign students individually in group assignments. When the team members develop the model together, it is quite difficult to accurately assess the individual contribution of each student. The individual submissions helped to encourage each student to engage in the modelling process. In addition, it helped to assess the individual component of the final score for the competition.

@@

Statistical Thinking (ETC2420/5242), covers regression, but not classification, and has a mix of undergraduate and postgraduate students. Only the 34 postgraduate (5242) students were required to participate in the kaggle competition, and competed in the regression (R) challenge. This was run independently from the MAST90083 competition. The 145 undergraduate (2420) students were used as comparison for examining performance of the postgraduate students. Although, it may be surprising at first glance, the undergraduate students provide a reasonable comparison for the graduate students. The class is taught to both cohorts simultaneously. The entry requirements to the Bachelors of Commerce at Monash is high, and these students have strong mathematics backgrounds. In the years prior to this experiment, the undergraduate scores on the final exam are indistinguishable from those of the graduate students. The competition ran for one month. Students formed their own teams of 2-4 members to compete. 
<!--
Several undergraduates also chose to compete individually. The material on regression methods, particularly the computational methods needed to successfully predict housing prices was new to both groups of students. 
-->
<!--
ETC3250, called Business Analytics, is an undergraduate course focusing on data mining. All students participated in a kaggle competition on the classification challenge. Because this group had no comparison group, it was difficult to assess performance. 
-->

@@

p1 <- ggplot(UniMelbData, aes(x=Proj_alloc, y=perf_class)) +
  geom_boxplot() +
  xlab("Classification competition") +
  ylab("Performance")
p2 <- ggplot(UniMelbData, aes(x=Proj_alloc, y=perf_reg)) +
  geom_boxplot() +
  xlab("Regression competition") +
  ylab("Performance")
grid.arrange(p1, p2, ncol=2)
```

@@

scores <- bind_rows(scores_5242, scores_2420)
ggplot(scores, aes(x=class, y=perf)) + 
  geom_boxplot() + 
  xlab("Participated in competition") +
  ylab("Performance")
```

>>>>

The reason for this strategy was first to motivate each of the students to think about modelling and be actively engaged in the competitions through individual submission. The lecturer allowed participants to create groups towards the end of the competition to illustrate the advantages of group work and ensemble models. Another reason for this approach was the university policy, requiring a strategy to assign students individually in group assignments. When the team members develop the model together, it is quite difficult to accurately assess the individual contribution of each student. The individual submissions helped to encourage each student to engage in the modelling process. In addition, it helped to assess the individual component of the final score for the competition.

@@

Table 1 shows some summary statistics for the two groups. We examine the percentage correct overall on the final exam for the different groups and the scores the students received for the second assignment. The second assignment examined students knowledge about computational methods and absolutely unrelated to the classification and regression methods. The difference between the two group are inconsiderable taking into the account the relatively small sample size of the groups.

<!-- The numbers calculated as part of the `Melb` chank -->

\begin{table}[h]
\begin{center}
\begin{tabular}{l l r r r r r r r}\hline
Score & Competition group & N & Mean (SD)   & Min  &  Q1 & Median  &  Q3  & Max \\\hline
Exam   & Regression      & 30 &  66.9 (14.1)  &   36 &  57.5 &   69.2 &  77.5 &   88.5\\
       & Classification & 31 &  62.1 (17.2)  &   20  & 52.2  &  68.0   &  75.2 &  85.5\\\hline

2$^{nd}$ assignment & Regression      &    30  & 8.16 (2.15)  &  0  & 7.02 &  8.35   &  9.9  &   10 \\
                    & Classification  &    31  & 8.01 (2.96)  &  0  & 7.8  &  9.30   &  9.9  &   10 \\\hline

\end{tabular}
\caption{MAST90083: Basics summary statistics of the exam score (out of 100) and the second assignment (out of 10) for the two competition groups.}
\end{center}
\label{tab:Melb_Group_Compare}
\end{table}

Statistical Thinking (ETC2420/5242), covers regression, but not classification, and has a mix of undergraduate and postgraduate students. Only the 34 postgraduate (5242) students were required to participate in the kaggle competition, and competed in the regression (R) challenge. This was run independently from the MAST90083 competition. The 145 undergraduate (2420) students were used as comparison for examining performance of the postgraduate students. Although, it may be surprising at first glance, the undergraduate students provide a reasonable comparison for the graduate students. The class is taught to both cohorts simultaneously. The entry requirements to the Bachelors of Commerce at Monash is high, and these students have strong mathematics backgrounds. In the years prior to this experiment, the undergraduate scores on the final exam are indistinguishable from those of the graduate students. The competition ran for one month. Students formed their own teams of 2-4 members to compete.

Similarly, to Table 1, Table 2 shows some summary statistics of the exam scores for the 34 postgraduate (ETC5242) students and for the 145 undergraduate (ETC2420) students. The mean and the median scores of postgraduate students are a bit lower than the corresponding scores of undergraduate students. However, the interquartile range is quite the same. Therefore, we do not have a principal evidence that postgraduate students are a stronger cohort. Interesting the highest exam score was received by an undergraduate student. Also, the mean and the median of the undergraduate students are a bit higher.

<!-- The numbers calculated as part of the `Melb` chank -->

\begin{table}[h]
\begin{center}
\begin{tabular}{l l r r r r r r r}\hline
Score & Competition group & N & Mean (SD)   & Min  &  Q1 & Median  &  Q3  & Max \\\hline
Exam   & ETC5242      & 34  & 61.8 (12.2)  &  35 & 57.2  &   61  & 72.2  & 81.5 \\
       & ETC2420         & 145 & 63.0 (15.4)  &   0 & 56.0  &   64  & 73.0   & 90.0\\\hline
\end{tabular}
\caption{ETC2420/5242: Basics summary statistics of the exam score (out of 100) for the two groups.}
\end{center}
\label{tab:Melb_Group_Compare}
\end{table}

Figure \ref{fig:Monash_quiz} presents the means and the standard deviations of the quizzes taken by the undergraduate (ETC2420) and postgraduate (ETC5242) students.

Postgraduate students have slightly higher average marks on quizzes and much smaller standard deviation. However, the median is only higher by up to 1 point for postgraduate students (except for quiz 5, the difference is 2 points). It could suggest that they are a bit stronger cohort than the undergraduate students. But it also could be that they have a more serious attitude to the quizzes. Considering their performance in the exam the latter is more likely.

```{r Monash_quiz, fig.cap = "\\label{fig:Monash_quiz} Means and error bars of undergraduate (ETC2420) and postgraduate (ETC5242) students quizes during the semester.", fig.height=4.5, fig.width=6}
library(tidyverse)
library(gridExtra)
quiz <- read_csv(file="data/quiz_scores_by_unit_2420_5442.csv")

quiz <- quiz %>% mutate(unit = as.character(unit))
quiz %>%ggplot(aes(x=unit, y=mean, colour=unit)) +
  geom_point(aes(color=unit), size = 2) +
  geom_errorbar(aes(ymin=mean-sd, ymax=mean+sd), width=.2, 
                position=position_dodge(.9)) +
  facet_wrap(~quiz)

# Calculate CI 
t <- quiz %>% select(unit,quiz,mean,sd) %>%
      gather(statistics, value, -(quiz:unit)) %>%
      unite(temp, unit, statistics) %>%
      spread(temp, value)
t <- t %>% mutate(mean_dif = `5242_mean` - `2420_mean`,
                  sd = sqrt(`5242_sd`^2/34 + `2420_sd`^2/145),
                  v1 = `5242_sd`^2/34,
                  v2 = `2420_sd`^2/145,
                  v = (v1+v2)^2/(v1^2/(34-1)  + v2^2/(145-1) ),
                  t_q = qt( .975, df=v),
                  lower.CI = mean_dif - t_q*sd,
                  upper.CI = mean_dif + t_q*sd
)   
```

<!--
Several undergraduates also chose to compete individually. The material on regression methods, particularly the computational methods needed to successfully predict housing prices was new to both groups of students. 
-->
<!--
ETC3250, called Business Analytics, is an undergraduate course focusing on data mining. All students participated in a kaggle competition on the classification challenge. Because this group had no comparison group, it was difficult to assess performance. 
-->

@@

p1 <- ggplot(UniMelbData, aes(x=Proj_alloc, y=perf_class)) +
  geom_boxplot() +
  xlab("Classification competition") +
  ylab("Performance")
p2 <- ggplot(UniMelbData, aes(x=Proj_alloc, y=perf_reg)) +
  geom_boxplot() +
  xlab("Regression competition") +
  ylab("Performance")
grid.arrange(p1, p2, ncol=2)

##### Summary statistics for comparing regression and classification groups:
Assig2_10p_sum <- UniMelbData %>% group_by(Proj_alloc)  %>%
      summarize( N=n(),
                     Mean = mean(Assig2_10p),
                     SD = sd(Assig2_10p),
                     Min = min(Assig2_10p),
                     Q1 = quantile(Assig2_10p, .25),
                     Median = quantile(Assig2_10p, .5), 
                     Q3 = quantile(Assig2_10p, .75),
                     Max = max(Assig2_10p)
             )

Exam_80_sum <- UniMelbData %>% group_by(Proj_alloc)  %>%
      summarize( N=n(),
                     Mean = mean(Exam_80),
                     SD = sd(Exam_80),
                     Min = min(Exam_80),
                     Q1 = quantile(Exam_80, .25),
                     Median = quantile(Exam_80, .5), 
                     Q3 = quantile(Exam_80, .75),
                     Max = max(Exam_80)
             )

@@

scores <- bind_rows(scores_5242, scores_2420)
ggplot(scores, aes(x=class, y=perf)) + 
  geom_boxplot() + 
  xlab("Participated in competition") +
  ylab("Performance")

scores_5242_exam <- scores_5242 %>%
                     summarize( 
                       N=n(),
                       Mean = mean(`Exam Tot`),
                       SD = sd(`Exam Tot`),
                       Min = min(`Exam Tot`),
                       Q1 = quantile(`Exam Tot`, .25),
                       Median = quantile(`Exam Tot`, .5), 
                       Q3 = quantile(`Exam Tot`, .75),
                       Max = max(`Exam Tot`))
scores_2420_exam <- scores_2420 %>%
                     summarize( 
                       N=n(),
                       Mean = mean(`Exam Tot`),
                       SD = sd(`Exam Tot`),
                       Min = min(`Exam Tot`),
                       Q1 = quantile(`Exam Tot`, .25),
                       Median = quantile(`Exam Tot`, .5), 
                       Q3 = quantile(`Exam Tot`, .75),
                       Max = max(`Exam Tot`))
```
